Reliable feature point matching

### 2013-08-15 text/head/abstracts.tex
Corrections

% English abstract
\cleardoublepage
\chapter*{Abstract}
%\markboth{Abstract}{Abstract}
\addcontentsline{toc}{chapter}{Abstract} % adds an entry to the table of 
% put your text here
Many algorithms have been proposed to solve the problem of matching 
feature points in two or more images using geometric assumptions to 
increase the robustness of the matching. However, very little work 
addresses the case where these assumptions might not hold. In particular 
few methods address the problem of reliable matching in cases where it 
is unknown whether two images have any corresponding areas or objects in 
the first place.  \\

We propose two algorithms for matching feature points without the use of 
geometric constraints. The first relies on the fact that any match 
between two images should be better than all possible matches within a 
single image. The second algorithm extends this idea by using community 
structure in the similarity graph of feature points to find reliable 
correspondences. To evaluate the algorithms experimentally, we introduce 
a simple method to generate a large amount of test cases based on a set 
of image pairs with viewpoint changes. Our results show that the 
proposed algorithm is generally superior to traditional approaches in 
finding correct correspondences.

>>>>

% English abstract
\cleardoublepage
\chapter*{Abstract}
%\markboth{Abstract}{Abstract}
\addcontentsline{toc}{chapter}{Abstract} % adds an entry to the table of 
% put your text here
Many algorithms have been proposed to solve the problem of matching 
feature points in two or more images using geometric assumptions to 
increase the robustness of the matching. However, very little work 
addresses the cases where these assumptions might not hold. In 
particular, few methods address the problem of reliable matching in 
cases where it is unknown whether two images have any corresponding 
areas or objects in the first place.  \\

We propose two algorithms for matching feature points without the use of 
geometric constraints. The first relies on the idea that any match 
between two images should be better than all possible matches within a 
single image. The second algorithm extends this idea by using the 
community structure in the similarity graph of feature points to find 
reliable correspondences. To evaluate the algorithms experimentally, we 
introduce a simple method to generate a large number of test cases based 
on a set of image pairs with viewpoint changes. Our results show that 
the proposed algorithm is generally superior to traditional approaches 
in finding correct correspondences.

### 2013-08-15 text/main/ch1_introduction.tex
Corrections

My goal when I started working on this thesis was not to improve feature 
point matching, but via a series of detours I ended up circling the 
problem which I found alluring, not just because it its importance, but 
also because of its inherent simplicity. Out of two sets, finding pairs 
of similar elements seems like a problem that should have been solved 
long ago, yet because of the nature of feature points and the data they 
are used on, there seems to be plenty of progress to be made still.

As the outcome of my work I propose two algorithms, \emph{Mirror Match 
(MM)} and \emph{Mirror Match with Clustering (MMC)}. Both algorithms are 
designed to tackle the problem of reliably matching feature points under 
circumstances where we cannot assume two images necessarily have any 
correspondences.  The goal of this thesis will be to describe these 
algorithms and convincingly demonstrate that they outperform traditional 
approaches.  While doing this I hope to provide sufficient background 
for a technically minded reader without any necessary experience in 
computer vision to follow the motivation behind \emph{MM} and \emph{MMC} 
and enable him or her to place both algorithms in a bigger scientific 
perspective.

The thesis is organized as follows: Chapter~\ref{C:Background} provides 
background and motivation for the development of \emph{MM} and 
\emph{MMC}. The chapter serves two purposes.  For those that might be 
reading this introduction wondering what a feature point really is and 
how it is matched, it provides a gentle introduction to the why's and 
how's of feature point matching. At the same time the chapter is meant 
as an introduction to related work which is hopefully useful for those 
who might use this thesis as a starting point for finding other research 
on the same subject. Chapter~\ref{C:Algorithms} will introduce in detail 
the algorithms proposed in this thesis. The reference algorithms used to 
benchmark \emph{MM} and \emph{MMC} are also introduced in detail.  
Chapter~\ref{C:Experiments} details the experiments done to evaluate the 
proposed algorithms both in terms of how they were conducted as well as 
their results. Chapter~\ref{C:Applications} provides two examples of how 
the matching algorithms can improve on panorama stitching and near 
duplicate matching.

For the remainder of the thesis I will make use of the academic 
'\emph{we}' to underline the fact this thesis had not been possible had 
it not been for the guidance of \emph{Stefan Winkler} and \emph{Sabine 
S\"usstrunk} as well as the support of \emph{Advanced Digital Science 
Center} in Singapore that hosted me as an intern during the work on this 
thesis.

>>>>

My original goal when I started working on this thesis was not to 
improve feature point matching. I started out working on face 
clustering, but via a series of detours I ended up circling the problem,
which I found alluring not just because it its importance, but also 
because of its inherent simplicity.  Out of two sets, finding pairs of 
similar elements seems like a problem that should have been solved long 
ago, yet because of the nature of feature points and the data they are 
used on, there still seems to be plenty of progress to be made.

As the outcome of my work I propose two algorithms, \emph{Mirror Match 
(MM)} and \emph{Mirror Match with Clustering (MMC)}. Both algorithms are 
designed to tackle the problem of reliably matching feature points under 
circumstances where we cannot assume two images necessarily have any 
correspondences.  The goal of this thesis will be to describe these 
algorithms and convincingly demonstrate that they outperform existing 
approaches.  While doing this I hope to provide sufficient background 
for a technically-minded reader without any necessary experience in 
computer vision to follow the motivation behind \emph{MM} and \emph{MMC} 
and enable him or her to understand both algorithms within a bigger 
scientific perspective.

The thesis is organized as follows: Chapter~\ref{C:Background} provides 
background and motivation for the development of \emph{MM} and 
\emph{MMC}. The chapter serves two purposes:  For those who might be 
reading this introduction wondering what a feature point really is and 
how it is matched, it provides a gentle introduction to the hows and 
whys of feature point matching. At the same time, the chapter is meant 
as an introduction to related work which is hopefully useful for those 
who might use this thesis as a starting point for finding other research 
on the same subject. Chapter~\ref{C:Algorithms} introduces in detail the 
algorithms proposed in this thesis. The reference algorithms used to 
benchmark \emph{MM} and \emph{MMC} are also introduced in detail.  
Chapter~\ref{C:Experiments} details both the setup and results of the 
experiments done to evaluate the proposed algorithms. 
Chapter~\ref{C:Applications} provides two examples of how the matching 
algorithms can improve on panorama stitching and near duplicate 
matching.

For the remainder of the thesis I will make use of the academic 
'\emph{we}' to underline the fact this thesis would not have been 
possible had it not been for the guidance of \emph{Stefan Winkler} and 
\emph{Sabine S\"usstrunk} as well as the support of \emph{Advanced 
Digital Science Center} in Singapore, which hosted me as an intern 
during the term in which I worked on this thesis.

### 2013-08-15 text/main/ch2_background.tex
Corrections

In this chapter we will introduce local feature points in images and 
explain how they are used. We will start out by discussing why we need 
local features in the first place before looking at the inner workings 
of modern local image features. Finally we will look at how they are 
used, in particular how corresponding pairs are found in between two or 
more images.

@@

The field of computer vision is anecdotally said to have been founded 
when Marvin Minsky assigned the problem of making a computer see as an 
undergraduate summer project back in 1966\footnote{This is the same 
Marvin who invented neural networks and went on to serve as scientific 
adviser on Kubrick's movie \emph{Space Odyssey 2001}}. Whether or not 
the anecdote is true, the idea that the field of computer vision should 
be solvable in a summer's worth of time but still to this days remains 
an active field of research, illustrates an interesting notion: Making 
computers solve tasks involving vision turns out to be much harder than 
we would initially assume.

Part of this notion stems from the fact that vision is a problem that we 
as humans are very good at. We can easily recognize objects under 
different light, stitch together photographies to a panorama or navigate 
a corridor identifying obstacles along the way. In fact these things are 
so easy to us, that it seems strange that they should be difficult at 
all, yet consistently recognizing an object under different lighting 
conditions and from different angles can still cause problems for even 
the best object recognition algorithms.

@@

The main factor that makes computer vision difficult is how the world we 
see around us constantly varies. This is not something that is 
immediately obvious to the casual observer, so for the sake of clarity 
let us do a small thought experiment. Try for a moment to imagine that 
you enter a room and recognize the newspaper that is lying on the table, 
illuminated from the grey light coming in of the window on the opposing 
wall. You move across the room and switch on the light make out if the 
paper is from today, and then you sit down and put on your glasses to 
read the headline. Try to imagine the newspaper as you see it at three 
different stages. First as you enter the slightly obscure room, then as 
you turn around after turning on the light, now standing on the other 
side of the table, and finally when you are sitting with the paper in 
your hands after you have put on your glasses. For each of these stages, 
try to mentally picture the situation as a photography and look at how 
the newspaper changes as you move around. At first you see it at an 
angle lying in a fairly dark and grey room, the white pages reflecting 
the light from the window with a cold bluish hue and the words too far 
away to make out.  Then in the second stage with the lights turned on, 
the object you are looking at has suddenly changed.  When you turned on 
the light in the room, you also changed the color of the newspaper, 
which now has a warm and yellowy white color to its pages.  The angle 
you are looking at the paper from has changed too, and the words and 
pictures are no longer upside down, although still slightly rotated and 
with the increased light in the room the contrast of the words and 
pictures of the page makes it possible for you to make recognize that 
the newspaper is from today.  Finally as you sit down and put your 
glasses on the newspaper changes again. The color might be the same, but 
if you imagine taking a photography as you are sitting with the paper in 
your hand, the newspaper will look much bigger, almost taking up the 
entire frame. Now that your glasses are on, the words stand out much 
clearer and you can see what is in the pictures on the front page.

Intuitively you know that the paper you sat down to read is the same 
object as the paper you saw when you entered the room, but if we limit 
ourselves to only looking at the object itself, almost everything about 
it changed in the process, from the color of the pages to the sharpness 
of the lines. On top of that, the paper has changed in size and rotation 
from lying on the table to being in your hand. In our daily lives we are 
aided in the process of recognizing the paper by our memory and idea of 
how the world works. When we went in to the dark room and saw something 
on the table, it was probably safe to assume that it was a newspaper.  
Maybe the newspaper always lies like that on the table, or maybe we left 
it there ourselves. Afterwards when we move around in the room, we can 
safely assume that the paper didn't move while our back is turned and 
even if it might not look exactly the same after we turned on the light, 
it's safe to guess that nobody took the old newspaper and replaced it 
with a new and yellower version.

@@

For this two work we need two components. First we need to find spots in 
an image that are distinct, since if we have similar points we can't 
uniquely match them. Secondly we need to take information around the 
local area of the point and describe it in a way that is invariant to 
changes like illumination and geometric distortion as well as noise, 
blur and scale. Finding unique points in the image is called 
\emph{feature detection} and it yields a \emph{key point}, where as we 
describe the key point with a \emph{feature descriptor}. The key point 
will contain the position of the point in the image while the descriptor 
will contain information about the image data surrounding that position.

@@

Intuitively the simplest way to find corners would be to detect edges in 
the image (for example using a high pass filter), and follow these edges 
until they turn. Some of these approaches are detailed in 
\cite{university1978comparison}. A different (and it turns out, better) 
approach is to look at the gradient across the image and find spots 
where there is a gradient in two perpendicular directions.  That is, 
along a dark edge on a bright background there will be a gradient 
perpendicular to the edge, but no gradient along the edge.  However in a 
corner we would have a gradient in two different directions. The idea 
was originally introduced by Morevec \cite{hans1977towards} and later 
improved upon by Harris and Stephens who introduced what is now commonly 
known as the Harris corner detector \cite{harris1988combined}.

Alternatively a fast way of deciding if a point is a corner is to look 
at all pixels surrounding the point of interest and compute the fraction 
of pixels that have higher intensities than the center\footnote{i.e.\ 
are brighter}. We can then find corners by thresholding on this 
intensity difference as proposed by Smith and Brady 
\cite{smith1997susan} as the SUSAN edge detector. Lepetit and Fua 
proposed a similar algorithm but only looking at the pixels on the 
circumference of a circle surrounding the point 
\cite{lepetit2006keypoint}. This is also the central idea behind the 
FAST detector by Rosten and Drummond \cite{rosten2006machine}.

@@

Once we have found a point of interest, we need to describe it. The 
simplest way is to save the values of the pixels in the immediate 
neighborhood, but this would not be very flexible. Take the case where 
we take two images of the same scene, but overexpose one of them. Two 
corresponding feature points would not look remotely alike in this case.
In the feature descriptor of bright photography all the pixel values 
would be high, while the values in the descriptor of the darker image 
would all be low. A similar situation arise if one image is rotated or 
there is a change in perspective.

A way to work around these issues is to compile a histogram of the 
pixels surrounding the key point. This way we would get the same 
descriptor now matter how the image is rotated. To a certain extend this
method is also invariant to affine transformations, that is change in 
perspective as long as approximately the same pixels fall within the 
zone we are taking a histogram of. If we use the raw values of the 
pixels, the descriptor is still not invariant with respect to 
brightness, but this is easily solved by creating a histogram of the 
pixel values relative to the center pixel. Lazebnik et al.\ introduced 
this method in \cite{lazebnik2003sparse} adding the distance from the 
key point as another dimension in the histogram.

Instead of using the pixel values, we can create a histogram of the 
gradients surrounding the feature point as Lowe did in his Scale 
Invariant Feature Detector (SIFT) \cite{lowe2004sift}. The histogram he 
proposes is based on the gradient locations and orientations weighted by
the gradient magnitude. To make sure that the histogram is rotation 
invariant, the bins of the histogram are normalized and start with the 
bin of the highest magnitude. The SIFT descriptor has been very widely 
used since its introduction and has become a de facto standard in 
computer vision. To enhance the computational time needed to compute 
descriptors, Bay et al.\ introduced the Speeded Up Robust Features 
(SURF) that rely on a histogram of wavelets instead of gradients as used 
in SIFT \cite{bay2006surf}.

A radically different approach to a feature descriptor is to measure the
intensity difference between a number of pixels surrounding the key 
point and simply assign the bits to a vector that is returned as the 
feature descriptor. Calonder et al.\ \cite{calonder2010brief} introduced 
a feature descriptor, Binary Robust Independent Elementary Features 
(BRIEF), based on this idea. The main advantage of their approach is 
reduced processing and memory needed to create, store and match the 
feature descriptors. Interestingly the positions used for pixel wise 
comparisons were selected at random and tested against similar random 
configurations. They then went on to use the configuration yielding the 
best result.  The simplicity of this method allows for real time 
matching of objects from a video stream, something that would be too 
computationally tolling when using a descriptor like SIFT\@. BRIEF 
however fails to be rotation invariant making it less useful in cases 
where images are not necessarily upright.

@@

Most applications of feature points use them to match features across 
different images. The standard way most feature points are matched 
involve comparing the feature descriptor using some kind of distance 
measure. If we match all features of one image with all features of 
another and discard all matches that are below a certain threshold of 
similarity, we have in essence a set of matches between the
two images. For most feature points (for example SIFT and SURF), this 
similarity is measured by the cosine similarity. Binary descriptors such
as BRIEF can take advantage of the faster computation of the hamming 
distance.

@@

While the standard matching method is often computationally efficient, 
it isn't necessarily the best way to match feature points. Many matching 
methods make use of the knowledge of the position of the keypoints in 
the image, i.e.\ the geometrical meta data. Using this information can 
often yield more accurate outcomes, but they inevitably make assumptions
about the geometric relationship between the images that are being 
matched which might not hold true.

A simple example would be to only consider matches that aren't deviating 
too much with respect to the average angle and distancen as considered 
by \cite{kim2008efficient}.  This performs well in situations where no 
camera rotation occurs between the two images, but fails when this 
assumption isn't met. Various scenarios have been proposed to improve on 
this simple assumption such as epipolar constraints 
(\cite{torr2000mlesac}, \cite{chum2005matching}) and pairwise 
constraints (\cite{choi2009robust}, \cite{leordeanu2005spectral} --- 
later referred to as \emph{Spectral}).  The Epipolar constraint carries 
the assumption that the two images matched are related by an affine 
transformation.  That is, that there is no relative movement of objects 
in between images and either the viewpoint is fixed or the image resides 
entirely on a plane.  In practice this assumption largely holds true 
when all objects we are interested in matching are roughly the same 
distance from the camera given that objects don't move across images.  
Pairwise constraints provide a more robust approach to the same problem 
by looking at a set of proposed correspondences and defining a pairwise 
error between any two matches for example based on the assumption that 
two neighboring correspondences will usually have similar angles and 
distances. This approach provides more robustness in cases where 
assumptions that are violated globally still hold on a local scale.

An alternative is to pick out different areas in each image and pair 
areas when correspondence pairs are mainly found going from one area to 
another. This allows for the filtering of all correspondences that do 
not fall within paired areas. Examples include \emph{Isomatch} 
\cite{das2008event}, which clusters feature points according to their 
position in the image, and the method from \cite{wu2011robust}, which 
uses maximally stable extremal regions (MSER) feature points to 
designate areas.

@@

In the absence of geometric constraints, only few methods are available.  
As pointed out in \cite{szeliski2010} and mentioned above, the simplest 
approach to matching feature points is to return all matches above a 
given threshold of similarity.  This can be done with a fixed threshold 
or by finding the nearest neighbors for each point (still filtering by a 
threshold to avoid including matches for feature points that have no 
counterpart in the other image).  When Lowe \cite{lowe2004sift} proposed 
the scale-invariant feature transform he included an alternative 
matching measure. Instead of rating matching by similarity assessed the 
uniqueness of a given match by looking at the two nearest neighbors of 
each feature point and calculating the matching score as the ratio of 
similarities. This method will later be referred to as \emph{Ratio}.  
Ranking the scores by their uniqueness and picking the $n$ best produces 
a set of correspondences that are distinctly matched across two images.  
Of these three methods (fixed threshold, nearest neighbors, and 
\emph{Ratio}), \emph{Ratio} was found to perform best by a slight margin 
\cite{mikolajczyk2005performance}.

@@

Based on this idea,  we determine reliable matches as follows. 
\emph{Mirror Match (MM)} matches every feature from both images with all 
other features and filters matches based on the ratio of best to second 
best match as in \emph{SIFT}. \emph{Mirror Match with Clustering (MMC)} 
on the other hand takes the combined set of feature points from both 
images and clusters these points according to their descriptors.  If a 
resulting partition contains only feature points from one image, no 
matches are returned.  If the partition contains points from both 
images, \emph{Mirror Match} is used to find the best matches within this 
partition.  The experimental results show that both approaches generally 
outperform existing correspondence matching methods.

>>>>

In this chapter we introduce the concept of local feature points in 
images and explain how they are used. We will start out by discussing 
the significance of local features in the first place before looking at 
the inner workings of modern local image features. Finally we will look 
at how they are used, in particular in terms of matching corresponding 
points between two or more images.

@@

The field of computer vision is anecdotally said to have been founded 
when Marvin Minsky assigned the task of making a computer see as an 
undergraduate summer project back in 1966\footnote{This is the same 
Marvin who invented neural networks and went on to serve as scientific 
adviser on Kubrick's movie \emph{Space Odyssey 2001}}.  Whether or not 
the anecdote is true, the fact remains that what was assigned as a 
summer's worth of work remains to this day an active field of research, 
which illustrates an interesting notion: Making computers solve tasks 
involving vision turns out to be much harder than we would initially 
assume.

Part of this notion of simplicity stems from the fact that vision is a 
task that we as humans are very good at. We can easily recognize objects 
under different light, stitch various perspectvies into a panorama or 
navigate a corridor identifying obstacles along the way.  In fact, these 
things are so easy to us, that it seems strange that they should be 
difficult at all; yet consistently recognizing an object under different 
lighting conditions and from different angles can still cause problems 
for even the best object recognition algorithms.

@@

The main factor that makes computer vision difficult is how the world we 
see around us constantly varies. This is not something that is 
immediately obvious to the casual observer, so for the sake of clarity 
let us do a small thought experiment. Try for a moment to imagine that 
you enter a room and recognize the newspaper that is lying on the table, 
illuminated from the grey light coming in of the window on the opposing 
wall. You move across the room and switch on the light make out if the 
paper is from today, and then you sit down and put on your glasses to 
read the headline. Try to imagine the newspaper as you see it at three 
different stages. First as you enter the slightly obscure room. Then as 
you turn around after turning on the light, now standing on the other 
side of the table; and finally when you are sitting with the paper in 
your hands after you have put on your glasses. For each of these stages, 
try to mentally picture the situation as a photography and look at how 
the newspaper changes as you move around. At first you see it at an 
angle lying in a fairly dark and grey room, the white pages reflecting 
the light from the window with a cold bluish hue and the words too far 
away to make out.  Then in the second stage with the lights turned on, 
the object you are looking at has suddenly changed.  When you turned on 
the light in the room, you also changed the color of the newspaper, 
which now has a warm and yellowy white color to its pages.  The angle 
you are looking at the paper from has changed too, and the words and 
pictures are no longer upside down, although still slightly rotated and 
with the increased light in the room the contrast of the words and 
pictures of the page makes it possible for you to make recognize that 
the newspaper is from today.  Finally as you sit down and put your 
glasses on the newspaper changes again. The color might be the same, but 
if you imagine taking a photography as you are sitting with the paper in 
your hand, the newspaper will look much bigger, almost taking up the 
entire frame. Now that your glasses are on, the words stand out much 
clearer and you can see what is in the pictures on the front page.

Intuitively you know that the paper you sat down to read is the same 
object as the paper you saw when you entered the room, but if we limit 
ourselves to only looking at the object itself, almost everything about 
it changed in the process, from the color of the pages to the sharpness 
of the lines. On top of that, the paper has changed in size and rotation 
from lying on the table to being in your hand.

In our daily lives we are aided in the process of recognizing the paper 
by our memory and idea of how the world works. When we went in to the 
dark room and saw something on the table, it was probably safe to assume 
that it was a newspaper.  Maybe the newspaper always lies like that on 
the table, or maybe we left it there ourselves. Afterwards when we move 
around in the room, we can safely assume that the paper didn't move 
while our back is turned and even if it might not look exactly the same 
after we turned on the light, it is reasonable to guess that nobody took 
the old newspaper and replaced it with a new and yellower version.

@@

For this to work we need two components. First we need to find spots in 
an image that are distinct, since if we have similar points we can't 
uniquely match them. Secondly we need to take information around the 
local area of the point and describe it in a way that is invariant to 
changes like illumination and geometric distortion as well as noise, 
blur and scale. Finding unique points in the image is called 
\emph{feature detection} and it yields a \emph{key point}, where as we 
describe the key point with a \emph{feature descriptor}. The key point 
will contain the position of the point in the image while the descriptor 
will contain information about the image data surrounding that position.

@@

Intuitively the simplest way to find corners would be to detect edges in 
the image (for example using a high-pass filter), and follow these edges 
until they turn. Some of these approaches are detailed in 
\cite{university1978comparison}. A different (and it turns out, better) 
approach is to look at the gradient across the image and find spots 
where there is a gradient in two perpendicular directions.  That is, 
along a dark edge on a bright background there will be a gradient 
perpendicular to the edge, but no gradient along the edge.  However in a 
corner we would have a gradient in two different directions. The idea 
was originally introduced by Morevec \cite{hans1977towards} and later 
improved upon by Harris and Stephens who introduced what is now commonly 
known as the Harris corner detector \cite{harris1988combined} building 
on the idea.

Alternatively a fast way of deciding if a point is a corner is to look 
at all pixels surrounding the point of interest and compute the fraction 
of pixels that have higher intensities than the center\footnote{i.e.\ 
are brighter}. We can then find corners by thresholding on this 
intensity difference as proposed by Smith and Brady 
\cite{smith1997susan} in their SUSAN edge detector. Lepetit and Fua 
proposed a similar algorithm which only took in to account the pixels on 
the circumference of a circle surrounding the point 
\cite{lepetit2006keypoint}. This is also the central idea behind the 
FAST detector by Rosten and Drummond \cite{rosten2006machine}.

@@

Once we have found a point of interest, we need to describe it. The 
simplest way is to save the values of the pixels in the immediate 
neighborhood, but this would not be very flexible. Take the case where 
we take two images of the same scene but overexpose one of them. Two 
corresponding feature points would not look remotely alike in this case.
In the feature descriptor of bright photography all the pixel values 
would be high, while the values in the descriptor of the darker image 
would all be low. A similar situation arise if one image is rotated or 
there is a change in perspective.

A way to work around these issues is to compile a histogram of the 
pixels surrounding the key point. This way we would get the same 
descriptor no matter how the image is rotated. To a certain extent this
method is also invariant to changes in perspective as long as 
approximately the same pixels fall within the zone we are taking a 
histogram of. If we use the raw values of the pixels, the descriptor is 
still not invariant with respect to brightness, but this is easily 
solved by creating a histogram of the pixel values relative to the pixel 
at the center of the feature point.  Lazebnik et al.\ introduced this 
method in \cite{lazebnik2003sparse} adding the distance from the key 
point as another dimension in the histogram.

Instead of using the pixel values, we can create a histogram of the 
gradients surrounding the feature point as Lowe did in his Scale 
Invariant Feature Detector (\emph{SIFT}) \cite{lowe2004sift}. The 
histogram he proposes is based on the gradient locations and 
orientations weighted by
the gradient magnitude. To make sure that the histogram is rotation 
invariant, the bins of the histogram are normalized and start with the 
bin of the highest magnitude. The \emph{SIFT} descriptor has been very 
widely used since its introduction and has become a de facto standard in 
computer vision. To enhance the computational time needed to compute 
descriptors, Bay et al.\ introduced the Speeded Up Robust Features 
(SURF) that rely on a histogram of wavelets instead of gradients as used 
in \emph{SIFT} \cite{bay2006surf}.

A radically different approach to a feature descriptor is to measure the
intensity difference between a number of pixels surrounding the key 
point and simply assign the bits to a vector that is returned as the 
feature descriptor. Calonder et al.\ \cite{calonder2010brief} introduced 
a feature descriptor, Binary Robust Independent Elementary Features 
(\emph{BRIEF}), based on this idea. The main advantage of their approach 
is the reduced processing and memory needed to create, store and match 
the feature descriptors. Interestingly the positions used for pixel wise 
comparisons were selected at random and tested against similar random 
configurations. They then went on to use the configuration yielding the 
best result as tested on a series of images. \emph{BRIEF} however is not 
rotation invariant making it less useful in cases where images are not 
necessarily upright.

@@

Most applications of feature points use them to match features across 
different images. The standard way most feature points are matched 
involve finding the distance between any two feature descriptors. If we 
match all features of one image with all features of another and discard 
all matches that are futher apart than some threshold distance, we have 
in essence a filtered set of matches between the two images. For most 
feature points (for example \emph{SIFT} and \emph{SURF}), this distance 
is measured by their cosine similarity. Binary descriptors such
as \emph{BRIEF} can take advantage of the faster computation of the 
hamming distance.

@@

While the standard matching method is often computationally efficient, 
it isn't necessarily the best way to match feature points. Many matching 
methods make use of the knowledge of the position of the keypoints in 
the image, i.e.\ the geometrical metadata. Using this information can 
often yield more accurate outcomes, but they inevitably make assumptions
about the geometric relationship between the images that are being 
matched which might not hold true.

A simple example would be to only consider matches that aren't deviating 
too much with respect to the average angle and distancen as considered 
by \cite{kim2008efficient}.  This method performs well in situations 
where no camera rotation occurs between the two images, but fails when 
this assumption isn't met. Various scenarios have been proposed to 
improve on this simple assumption such as epipolar constraints 
(\cite{torr2000mlesac}, \cite{chum2005matching}) and pairwise 
constraints (\cite{choi2009robust}, \cite{leordeanu2005spectral} --- 
later referred to as \emph{Spectral}).  The Epipolar constraint carries 
the assumption that the two images matched are related by an affine 
transformation.  That is; there is no relative movement of objects in 
between images and either the viewpoint is fixed or the image resides 
entirely on a plane.  In practice this assumption largely holds true 
when all objects we are interested in matching are roughly at the same 
distance from the camera given that objects stay fixed from one image to 
another.  Pairwise constraints provide a more robust approach to the 
same problem by looking at a set of proposed correspondences and 
defining a pairwise error between any two matches for example based on 
the assumption that two neighboring correspondences will usually have 
similar angles and distances. This approach provides more robustness in 
cases where assumptions that are violated globally still hold on a local 
scale.

An alternative is to pick out different areas in each image and pair 
areas when correspondence pairs are mainly found going from one area to 
another. This allows for the filtering of all correspondences that do 
not fall within paired areas. Examples include \emph{Isomatch} 
\cite{das2008event}, which clusters feature points according to their 
position in the image, and the matching method from \cite{wu2011robust}, 
which uses Maximally Stable Extremal Regions (\emph{MSER}) feature 
points to designate areas.

@@

In the absence of geometric constraints, only few methods are available.  
As pointed out in \cite{szeliski2010} and mentioned above, the simplest 
approach to matching feature points is to return all matches above a 
given threshold of similarity.  This can be done with a fixed threshold 
or by finding the nearest neighbors for each point while still filtering 
by a threshold to avoid including matches for feature points that have 
no counterpart in the other image.  When Lowe \cite{lowe2004sift} 
proposed the scale-invariant feature transform he included an 
alternative matching measure. Instead of rating matching by similarity 
he assessed the uniqueness of a given match by looking at the two 
nearest neighbors of each feature point and calculating the matching 
score as the ratio of similarities. This method will later be referred 
to as \emph{Ratio}.  Lowe ranks the scores by their uniqueness and picks 
the $n$ best ones, which yields a set of correspondences that are 
distinctly matched across two images.  Of these three methods (fixed 
threshold, nearest neighbors, and \emph{Ratio}), \emph{Ratio} was found 
to perform best by a slight margin \cite{mikolajczyk2005performance}.

@@

Based on this idea,  we determine reliable matches as follows. 
\emph{Mirror Match (MM)} matches every feature from both images with all 
other features and filters matches based on the ratio of best to second 
best match as in \emph{SIFT}. \emph{Mirror Match with Clustering (MMC)} 
on the other hand takes the combined set of feature points from both 
images and clusters these points according to their descriptors.  If a 
resulting partition contains only feature points from one image, no 
matches are returned.  If the partition contains points from both 
images, \emph{Mirror Match} is used to find the best matches within this 
partition.  The experimental results show that both approaches generally 
outperform existing matching methods.
